### 2019-11-29 manuscript/field_data.tex

\section{Field data test}
Test with real data are conducted with the gravity data from Caraj\'as, north of Brazil, were provided by the  Geological 
Survey of Brazil (CPRM). The real aerogravimetric data were collected in 113 flight lines along north–south direction 
with flight line spacing  of 3 km and tie lines along east–west direction at 12 km.

This airborne gravity survey was divided in two different areas, collected in different times, having samples spacing of 
$7.65$ m and $15.21$ m, totalizing  $4\,353\,428$ observation points. The height of the flight was fixed at 900 m. 
The gravity data (Figure \ref{fig:carajas_real_data}) were gridded into a regularly spaced dataset of $250\,000$ 
observation points ($500 \times 500$) with a grid spacing of $716.9311$ km north-south and $781.7387$ km east-west.

To apply our modification of the fast equivalent-layer method \cite[]{siqueira-etal2017}  that computes the forward 
modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}), we set an equivalent layer at 300 m deep. 
Figure \ref{fig:carajas_gz_predito_val}a shows the fitted gravity data after 50 iterations by applying our method. 
The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed 
(Figure \ref{fig:carajas_real_data}) and the predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show an 
acceptable data fitting because they have a mean close to zero ($0.0003$ mGal) and a small standard deviation of 
$0.105$ mGal which corresponds to approximately $0.1$ \% of the amplitude of the gravity data.

These small residuals indicate that our method yielded an estimated mass distribution (not shown) that can be used in 
the data processing. We perform upward-continuation of the real gravity data (Figure \ref{fig:carajas_real_data}) at 
a constant height of $5\,000$ m over the real data. The upward-continued gravity data 
(Figure \ref{fig:up2000_carajas_500x500}) seem a reasonable processing because of the attenuation of the short 
wavelenghts. By using our approach, the processing of the $250\,000$ observations was extremely fast and took 
$0.216$ seconds.