### 2019-11-29 manuscript/manuscript.tex

%\documentclass[paper,twocolumn,twoside]{geophysics}
\documentclass[manuscript,blind]{geophysics}
%\documentclass[manuscript]{geophysics}

% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

% Extra packages
\usepackage{amsmath}
\usepackage{bm}
\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
	allcolors=black,
}
\usepackage{lipsum}
\usepackage[table]{xcolor}

\begin{document}

\title{Fast equivalent-layer technique for gravity data processing with BTTB matrix systems}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-XXXX} % manuscript number

%\address{
%\footnotemark[1]Observat\'{o}rio Nacional, \\
%Department of Geophysics, \\
%Rio de Janeiro, Brazil}
%\author{Diego Takahashi\footnotemark[1], Vanderlei C. Oliveira{ }Jr.\footnotemark[1] and 
%Val{\'e}ria C. F. Barbosa\footnotemark[1]}
%
%\footer{vanderlei@on.br}
%\lefthead{Takahashi, Oliveira{ }Jr. \& Barbosa}
%\righthead{Fast equivalent-layer for BTTB systems}

\maketitle

% Main body
\include{abstract}
\include{introduction}
\include{methodology}
\include{synthetic_data}
\include{field_data}
\include{conclusions}
\include{aknowledgements}

% Tables and figures
\include{tables}
\include{figures}

% Appendices
\include{DFT_computations}
\include{eigenvalues_C}

\bibliographystyle{seg}  % style file is seg.bst
\bibliography{references}

\end{document}