### 2019-11-29 manuscript/tables.tex

\tabl{RAM-usage}{Comparison between the system RAM memory usage needed to store the full 
matrix $\mathbf{A}$ (equations \ref{eq:predicted-data-vector} and \ref{eq:BTTB_A}), a single 
column of $\mathbf{A}$ and the full matrix $\mathbf{L}$ (equation \ref{eq:DTF_G}). The quantities 
are represented in megabyte (MB), for different total number of observations $N$. We consider 
that the elements of $\mathbf{A}$ and $\mathbf{L}$ occupy, respectively, 8 and 16 bytes each 
in computer memory.
\label{tab:RAM-usage}
}{
	\begin{center}
		\begin{tabular}[]{|l|c|c|c|}
			\hline
			\textbf{$N$} & \textbf{Full matrix $\mathbf{A}$} & \textbf{Single column of $\mathbf{A}$} 
			& \textbf{Full matrix $\mathbf{L}$}\\
			\hline 
			$100$ & 0.0763 & 0.0000763 & 0.0006104\\
			\hline
			$400$ & 1.22 & 0.0031 & 0.0248\\
			\hline
			$2\,500$ & 48 & 0.0191 & 0.1528\\
			\hline
			$10\,000$ & 763 & 0.00763 & 0.6104\\
			\hline
			$40\,000$ & 12\,207 & 0.305 & 2.4416 \\
			\hline
			$250\,000$ & 476\,837 & 1.907 & 15.3 \\
			\hline
			$500\,000$ & 1\,907\,349 & 3.815 & 30.518 \\
			\hline
			$1\,000\,000$ & 7\,629\,395 & 7.629 & 61.035 \\
			\hline
		\end{tabular}
	\end{center} 
}