### 2020-03-21 manuscript/manuscript.tex

%\documentclass[paper,twocolumn,twoside]{geophysics}
\documentclass[manuscript,noblind]{geophysics}
%\documentclass[manuscript]{geophysics}

% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

% Extra packages
\usepackage{amsmath}
\usepackage{bm}
\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
	allcolors=black,
}
\usepackage{lipsum}
\usepackage[table]{xcolor}

\begin{document}

\title{Convolutional equivalent layer for gravity data processing}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-XXXX} % manuscript number

\address{
\footnotemark[2]Observat\'{o}rio Nacional, Department of Geophysics, Rio de Janeiro, Brazil\\
\footnotemark[1] Corresponding author: vanderlei@on.br
}
\author{Diego Takahashi\footnotemark[2], Vanderlei C. Oliveira{ }Jr.\footnotemark[2]\footnotemark[1] and 
Val{\'e}ria C. F. Barbosa\footnotemark[2]}

%\footer{Example}
%\lefthead{Takahashi, Oliveira{ }Jr. \& Barbosa}
\righthead{Convolutional equivalent layer}

\maketitle

% Main body
\include{abstract}
\include{introduction}
\include{methodology}
\include{synthetic_data}
\include{field_data}
\include{conclusions}
%\include{aknowledgements}

% Tables and figures
\include{tables}
\include{figures}

% Appendices
\include{DFT_computations}
\include{eigenvalues_C}

\bibliographystyle{seg}  % style file is seg.bst
\bibliography{references}

\end{document}