### 2020-03-26 manuscript/figures.tex
include new figures and modifications in text

\renewcommand{\figdir}{Fig} % figure directory

\plot{Figure1}{width=\textwidth}{
	{Schematic representation of an $N_{x} \times N_{y}$ regular grid of points (black dots) defined by 
	$N_{x} = 4$ and $N_{y} = 3$. The grids are oriented along the (a) $x$-axis and (b) $y$-axis. The grid 
	coordinates are $x_{k}$ and $y_{k}$, where the $k = 1, \dots, N_{x}$ and $l = 1, \dots, N_{y}$ are 
	called the grid indices. The insets show the grid indices $k$ and $l$.}
	\label{fig:methodology}
}

% Computational performance

\plot{Figure2}{width=\textwidth}{
	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 
	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ 
	iterations and $\kappa = 5$.}
	\label{fig:float}
}

\plot{Figure3}{width=10cm}{
	{Comparison between the running time of the fast equivalent-layer technique \citep{siqueira-etal2017} and our method.
	The values were obtained for $N^{it} = 50$ iterations.}
	\label{fig:time_fast_eqlayer_bccb}
}

\plot{Figure4}{width=10cm}{
	{Running time of our method for a number of observations $N$ up to $25$ millions. 
	The values were obtained for $N^{it} = 50$ iterations.}
	\label{fig:time_bccb}
}

% Application to synthetic data

\plot{Figure5}{width=13.5cm}{
	{Application to synthetic data. 
	Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
	a sphere with density contrast $-0.5\, \mathrm{g/cm^3}$ and 
	two rectangular prisms with density contrasts $0.35\, \mathrm{g/cm^3}$ (upper-left body) and 
	$0.4\, \mathrm{g/cm^3}$ (upper-right body). The black lines represent the horizontal projection 
	of the sources.}
	\label{fig:synthetic_data}
}

\plot{Figure678_r1}{width=6cm}{
	{Application to synthetic data. 
	(a) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
	predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
	The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal, respectively.
	(b) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
	predicted data produced by our method. The mean and standard deviation are 
	$8.264 \times 10^{-7}$ and $0.0144$ mGal, respectively.
	(c) Difference between the predicted data produced by the fast equivalent-layer technique 
	and our method.}
	\label{fig:comparison_datafits}
}

\plot{Figure10_r1}{width=6cm}{
	{Application to synthetic data.
	(a) Gravity data produced by the synthetic sources at $z = -XXXX$ m.
	(b) Residuals between the simulated data shown in a and the upward-continued 
	gravity data obtained with the fast equivalent-layer technique (not shown), with 
	mean $XXXXXX$ mGal and standard deviation $XXXXXX$ mGal. 
	(c) Residuals between the simulated data shown in a and the upward-continued 
	gravity data obtained with our method (not shown), with mean $XXXXX$ mGal and 
	standard deviation $XXXXX$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and 
	our method were $7.62026$ and $0.00834$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{Figure11_r1}{width=6cm}{
	{Application to synthetic data. 
	(a) Gravity data produced by the synthetic sources at $z = -XXXX$ m.
	(b) Residuals between the simulated data shown in a and the downward-continued 
	gravity data obtained with the fast equivalent-layer technique (not shown), with 
	mean $XXXXXX$ mGal and standard deviation $XXXXXX$ mGal. 
	(c) Residuals between the simulated data shown in a and the downward-continued 
	gravity data obtained with our method (not shown), with mean $XXXXX$ mGal and 
	standard deviation $XXXXX$ mGal. 
	The total computation times spent by the fast equivalent-layer technique and 
	our method were $7.59654$ and $0.00547$ seconds, respectively.}
	\label{fig:downward_comparison}
}

% Application to field data

\plot{Figure12}{width=13.5cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	Observed gravity data on a regular grid of $500 \times 500$ points, totaling $N = 250,000$ 
	observations. The inset shows the study area (blue rectangle) which covers the southeast part of 
	the state of Par{\'a}, north of Brazil.}
	\label{fig:carajas_real_data}
}

\plot{Figure13_r1}{width=8cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	(a) Predicted data produced by our method. (b) Residuals between the observed  
	(Figure \ref{fig:carajas_real_data}) and the predicted data (panel a), with mean 
	$0.000292$ mGal and standard deviation of $0.105$ mGal.}
	\label{fig:carajas_gz_predito_val}
}

\plot{Figure14}{width=13.5cm}{
	{Application to field data over the Caraj{\'a}s Province, Brazil. 
	The upward-continued gravity data obtained with our method $5\,000$ m above the 
	observed data (Figure \ref{fig:carajas_real_data}).
	The total computation time for processing of the $250,000$ observations was $0.216$ seconds.}
	\label{fig:up2000_carajas_500x500}
}