### 2020-04-30 manuscript/abstract.tex
include the classical FFT results in abstract

\begin{abstract}

We have developed an efficient and very fast equivalent-layer technique for gravity data processing by modifying an 
iterative method grounded on excess mass constraint that does not require the solution of linear systems. 
Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure of the sensitivity matrix, that raises 
when regular grids of observation points and equivalent sources (point masses) are used to set up a fictitious 
equivalent layer, we have developed an algorithm which greatly reduces the number of flops and RAM memory necessary 
to estimate a 2D mass distribution over the equivalent layer. The structure of symmetric BTTB matrix consists of 
the elements of the first column of the sensitivity matrix, which in turn can be embedded into a symmetric 
Block-Circulant Circulant-Block (BCCB) matrix. Likewise, only the first column of the BCCB matrix is needed 
to reconstruct the full sensitivity matrix completely. From the first column of BCCB matrix, its eigenvalues 
can be calculated using the 2D Fast Fourier Transform (2D FFT), which can be used to readily 
compute the matrix-vector product of the forward modeling in the fast equivalent-layer technique. As a result, 
our method is efficient to process very large datasets using either fine- or mid-grid meshes. The larger the 
dataset, the faster and more efficient our method becomes compared to the available equivalent-layer techniques. 
Tests with synthetic data demonstrate the ability of our method to satisfactorily upward-
and downward-continuing the gravity data.
Our results show very small border effects and noise amplification compared to those 
produced by the classical approach in the Fourier domain.
Besides, they show that while the running time of our method is $\approx 30.9$ seconds for 
processing $N = 1\,000\,000$ observations, the fast equivalent-layer technique spent 
$\approx 46.8$ seconds with $N = 22\,500$.
A test with field data from Caraj{\'a}s province, Brazil, illustrates the low computational 
cost of our method to process a large data set composed of $N = 250\,000$ observations.

\end{abstract}